\chapter{Introduction}
\label{chp:intro}

\section{Unsafe Code}
Browsers and the Internet-of-Things (IoT) require running untrusted code (\ie unknown code from an unknown source that could do anything), that may have been downloaded from anywhere.
It is crucial to ensure the safety of the code being executed in these contexts.
Otherwise, one website may crash your entire browser/IoT device, read secret data in your browser/IoT device, or attempt to take over control browser/IoT device.
Many such exploits have originated due to unsafe code being downloaded and run.
Typically, \emph{sandboxing} and/or \emph{dynamic safety checks} are used to ensure the safety of untrusted code.

Sandboxing involves placing untrusted code into a secure environment to contain the damage caused by unsafe behavior \cite{sandboxes}.
For example, Google Chrome places the running scripts for each website in separate processes so that unsafe code cannot access the address space of other websites or the browser \cite{chromesandbox}.
This prevents websites from crashing, stealing data, or taking control of other websites and the browser.
However, sandboxing in this way requires more run time resources than running scripts in one process, as processes require overhead in most OSes.

Dynamic safety checks are run time checks that catch any attempted unsafe operations.
For example, WebAssembly (\wasm) is a low-level language designed to be both safe and fast to execute in place of JavaScript for performance-critical applications in browsers.
While \wasm is type safe, memory safe, and its semantics enforce the separation of control flow and data (which together work like sandboxing to, \eg prevent websites from crashing each other), it still relies on dynamic checks to ensure these properties at run time.
These dynamic checks potentially slow down programs by introducing unnecessary instructions to perform the checks.

We have designed an extension to \wasm, called \name, that adds new instructions that do not require dynamic safety checks.
However, under the existing \wasm model the new \name instructions have potentially unsafe semantics, as they require stronger static guarantees than \wasm can provide to ensure the same level of safety as \wasm.
These instructions are likely to be faster than their \wasm counterparts because they do not require the addition of instructions by the compiler/interpreter to perform checks.
To provide these additional static guarantees, we equip \name with a more advanced type system than \wasm.

\section{Type Systems}
Types systems are useful for reasoning about programs.
They can be used to reason about the behavior of programs, usually in the form of safety guarantees.
For example, type safety is the property that a well-typed program will never become \emph{stuck}; that is, an expression will always be a well-formed irreducible value, reduce to an error, or reduce to another well-formed expression.
The safety guarantees of type systems provide a degree of trust in programs, as a well-typed program implicitly contains a checkable proof that it will only exhibit well-defined behavior (which, in the case of \wasm, has several safety guarantees as discussed above).

Generally, low-level languages are either untyped or have simple type systems that provide minimal guarantees.
More expressive type systems can encode richer invariants, enabling ruling out more bad behaviors with static checks alone.
Generally, such type systems are attached to high-level languages, where explicit abstractions make it easier to reason about programs compared to low-level languages.
Conversely, using expressive type systems in low-level languages often requires reasoning about program state and unstructured control flow (\eg \texttt{goto}), which introduces more complexity into the type system.
However, prior work has attached expressive type systems that permit rich correctness guarantees to simple low-level languages.

\section{Related Work}
\label{sec:relwork}
%% TIL
Using type information to improve compiler optimizations is not a new idea.
\citet{TIL} used strongly typed intermediate languages (TIL) to improve optimizations of SML code.
Compiling SML involves many translations among intermediate languages, and by preserving type information across those translations \citet{TIL} were able to safely perform additional compiler operations.
Using TIL in the compilation of programs led to up to $50\%$ faster programs.
TIL focuses on compiler optimizations and eventually translates into untyped languages and finally runnable assembly, so the ensured guarantees of the type system are lost along the way (they may still be present, but it is no longer possible to statically check them).

%% PCC
Proof-carrying code (PCC) was introduced by \citet{PCC}.
PCC attached explicit proofs that low-level code satisfies some safety properties.
The proof can then be checked to ensure the safety of the code before it is run.
While typed intermediate languages require types to be considered as part of the language, PCC adds a separate logical framework on top.
Thus, PCC can be used with any language, and requires no support from the language.
However, because PCC has no support from the language, it has a higher proof burden than using types inside a language.

\citet{PCC} provides a detailed example of invariants for extensions to TIL to ensure type safety of compiled code.
The example uses the Edinburgh Logical Framework (LF) to encode the proof.
A type safety proof of a LF program is a proof of correctness (with respect to some specification).
Encoded proofs may be quite large, requiring extra time to transmit and check compared to type annotations and typechecking.

%% FToTAL
\citet{FtoTAL} showed how types could provably be preserved during five different compilation passes to get from System F all the way down to a typed assembly language (TAL).
The purpose of TAL was focused on safety.
Although \citet{FtoTAL} argued that the type-preserving compilation passes would permit similar optimizations to TIL, they didn't include further optimizations based on TAL.
However, \citet{FtoTAL} did argue that the guarantees of TAL were sufficient to allow untrusted code to be safely executed.

%% DTAL
\citet{DTAL} created a much more expressive type system for an assembly language which had the potential to allow more compiler optimizations.
Their language, a dependently typed assembly language (\dtal), used a limited dependent type system, which enabled safely removing some run time checks, including array bounds checks.
The goal of \dtal, similar to TAL, was to support type-preserving compilation from a high-level language for both optimizations and safety.
\dtal intended to support type-preserving compilation from Dependent ML as well as SML.

%% Wasm
After almost two decades of JavaScript being the dominant language in browsers, it was decided that an alternative was necessary for performance-critical code.
The alternative that was jointly created by the major browser developers was WebAssembly (\wasm) \cite{WASM}.
\wasm is a stack-based assembly language with structured control flow.
It is designed to be safe as well as performant, with a small binary footprint.
The \wasm type system is simple, only encoding primitive types, but strong enough to ensure certain safety properties (\ie that arbitrary code cannot be executed).
Memory safety in \wasm is enforced using run time checks.
\wasm is supported by most major browsers, and can work well for IoT devices due to its portability and safety.

\section{Contributions}
We want to use types to improve performance while ensuring safety in real-world low-level programs.
While prior work has used expressive type systems for low-level languages, we want to show the feasibility of doing this with a language commonly used in contexts that are both performance-critical and untrusted.
Towards that goal, we introduce \name, an extension of the WebAssembly (\wasm) language.

We chose \wasm to build on because it is used in browsers and IoT devices.
Both browsers and IoT device require strong safety guarantees (such as preventing programs from crashing the whole browser/IoT device) since they download and execute code from unknown sources.
Further, the use case for \wasm is often performance-critical applications such as in-browser games and cryptographic algorithms.

\name introduces new versions of \wasm instructions which require no dynamic checks, but also require stronger type-level safety guarantees than their \wasm counterparts (\autoref{sec:newinstructions}).
To facilitate type-checking these new instructions, \name uses an indexed type system which is able to encode linear constraints on program variables and therefore ensure safety properties of the new \name instructions (\autoref{sec:typesys}).
We ensure that \name is as safe as \wasm by providing a type safety proof of the \name indexed type system (\autoref{sec:typesafety}).
Together, these additions mean that \name is as safe as \wasm while potentially improving performance by moving run-time checks to compile time.

We show that \name is backwards compatible with \wasm by showing that \wasm programs can be turned into \name programs (\autoref{subsec:embedding}).
This means that there is no additional proof burden on the programmer to use \name.
We also show that we can go the other way by turning \name programs into well-typed \wasm programs (\autoref{subsec:erasure}).

\section{Thesis Statement}
\begin{adjustwidth}{1cm}{1cm}
    An indexed type system can be used in a low-level language to reduce the number of dynamic checks required, without sacrificing safety and security guarantees or increasing the programmer's proof burden.
\end{adjustwidth}
