\chapter{Metatheory}
\label{chp:metatheory}

Now that we have introduced \name and shown how it can be used for reasoning, it is time to reason about \name itself.
First, we will take a look at the relationship between \wasm and \name, by showing methods to convert \wasm programs to \name programs and vice versa.
Then, we will prove the type safety of \name, to ensure that our claim that \name is as safe as \wasm is valid.
However, before we can do any of that, we must ``complete'' our reasoning ability by creating a way to connect the reduction relation form with the type system.

\section{Administrative Typing Rules}
While we have shown the \name typing rules for instructions within a static context, we still need typing rules for administrative instructions and the store used in reduction.
\emph{Administrative instructions} are introduced for reduction to keep track of information during reduction.
For example, $\<local>$ is the result of reducing a closure call; it is used to reduce a function body within the closed environment of the closure.
They are not part of the surface syntax of a language (\eg you cannot put a local block in a \name program), and can only appear as an intermediate term during reduction.
\autoref{fig:storerules} shows the \name typing rules for module instances $inst$, the run time store $s$, and various data structures contained within $s$.
There are many different judgments being introduced, so we explicitly state the form of the judgment before stating the rule for that judgment.

During reduction, we use \refrule{Program} (\autoref{fig:programrules}) to ensure that a \name program (consisting of the store $s$, local variables $v^{*}$, and instruction sequence $e^{*}$) is well typed (notice that it has the same form as the reduction relation).
It uses \refrule{Code} and \refrule{Store}, to ensure that a reducible \name program is well typed.
\refrule{Code} checks that a sequence of instructions is well typed with an empty stack, the indexed types and constraints for the given local variables in the precondition, and an optional return postcondition (not used by \refrule{Program}).
Since local variables are values, we know that each one of them is equal to some constant, so \refrule{Code} is really just checking that the sequence of instructions has some postcondition reachable from the given local variables.
There is an optional return postcondition for \refrule{Code} because \refrule{Local} has as a premise a judgment of the exactly same form, except with a return postcondition.

\begin{figure}
    $$ S ::= \{ \text{inst } C^{*}, \text{ tab } n^{*}, \text{ mem } m^{*} \} $$

    $ \boxed{\vdash s;v^{*};e^{*}} $

    \begin{mathpar}
        \inferrule*[right=\defrule{Program}]{ %% admin program
            \vdash s : S \and
            S;\epsilon \vdash_i v^{*};e^{*} : ti^{*};l;\phi
        } {
            \vdash_i s;v^{*};e^{*} : ti^{*};l;\phi
        }
    \end{mathpar}

    $ \boxed{S;(ti^{*};l;\phi)^{?} \vdash_i v^{*};e^{*} : ti^{*};l;\phi} $

    \begin{mathpar}
    \inferrule*[right=\defrule{Code}]{ %% admin code
        (\vdash v : \ti{t}{a};\phi_v)^{*}\\
        C = S_{\text{inst}}(i),\text{local} \; t^{*}, \text{return} \; (ti^n;l;\phi)^{?}\\
        S;C \vdash e^{*} : \epsilon:\ti{t}{a}^{*};\phi_v^{*} \rightarrow ti^n;l;\phi
    } {
        S;(ti^n;l;\phi)^{?} \vdash_i v^{*};e^{*} : ti^n;l;\phi
    }
    \end{mathpar}
    \caption{\name Program Typing Rules}
    \label{fig:programrules}
\end{figure}

In addition to getting the type of the instructions being reduced, we also need to know the type of the store $s$ since it is part of the reduction relation.
\refrule{Store} checks that a run-time store, $s$ is well typed by the store context $S$.
The store context $S$ is to $s$ as $C$ is to $inst$.
That is, it contains the type information for everything in $s$.
\refrule{Store} ensures that every module instance $inst$ in $s$ has the type of the index module context $C$ in $S$ using \refrule{Instance}.
Further, \refrule{Store} ensures that all of the closures in all of the tables in $s$ are well typed, and the the sizes of all the tables and memory chunks in $S$ do not exceed the actual size of their implementations.

To get the type of the store, we in turn have to know the types of each of the various run-time data structures.
\refrule{Instance} checks that a module instance is well-typed by the index module context under the store context $S$.
It checks all of the closures $cl^{*}$ against their expected types $tfi^{*}$ in $C$, and similarly for all of the globals ($v^{*}$ and $()\text{mut}^{?} t)^{*}$).
The table and memory indices ($i$ and $j$, respectively) are used to lookup the the relevant types ($(n,tfi^{*})$ and $m$, respectively) in the store context $S$.
Closures are typechecked by \refrule{Closure}, which falls back on the module typing rules from \autoref{fig:modulerules} to typecheck the function definition inside of the closure.
\refrule{Admin-Const} gets the postcondition indexed types and constraints on values; it is used to typecheck local and global variables.

\begin{figure}
    $ \boxed{\vdash s : S} $

    \begin{mathpar}
        \inferrule*[right=\defrule{Store}]{ %% store
            S = \{ \text{inst} \; C^{*}, \text{tab} \; n^{*}, \text{mem} \; m^{*} \}\\
            (S \vdash inst : C)^{*} \and
            ((S \vdash cl : tfi)^{*})^{*} \\
            (n \leq |cl^{*}|)^{*} \and
            (m \leq |b^{*}|)^{*}
        } {
            \vdash \{ \text{inst} \; inst^{*}, \text{tab} \; (cl^{*})^{*}, \text{mem} \; (b^{*})^{*} \} : S
        }
    \end{mathpar}

    $ \boxed{S \vdash inst : C} $

    \begin{mathpar}
        \inferrule*[right=\defrule{Instance}]{ %% instance
            (S \vdash cl : tfi)^{*} \and
            (\vdash v : \ti{t}{a},\phi_v)^{*} \\
            (S_\text{tab}(i) = n)^{?} \and
            (S_\text{mem}(j) = m)^{?}
        } {
            S \vdash
            {\begin{stackTL}
                \{ \text{func} \; cl^{*}, \text{glob} \; v^{*}, \text{tab} \; i^{?}, \text{mem} \; j^{?} \}
                \\ : \{ \text{func} \; tfi^{*}, \text{global} \; (\text{mut}^{?} \; t)^{*}, \text{table} \; n^{?}, \text{memory} \; m^{?} \}
            \end{stackTL}}
        }
    \end{mathpar}

    $ \boxed{\vdash v : ti;\phi} $

    \begin{mathpar}
        \inferrule*[right=\defrule{Admin-Const}]{ %% admin const
        } {
            \vdash t.\<const> c : \ti{t}{a};\circ,\ti{t}{a},(\<eq> a \; \ti{t}{c})
        }
    \end{mathpar}

    $ \boxed{S \vdash cl : tfi} $

    \begin{mathpar}
        \inferrule*[right=\defrule{Closure}]{ %% closure
            S_\text{inst}(i) \vdash f : tfi
        } {
            S \vdash \{ \text{inst} \; i, \text{code} \; f \} : tfi
        }
    \end{mathpar}
    \caption{\name Store Typing Rules}
    \label{fig:storerules}
\end{figure}

\begin{figure}
    $\boxed{S;C \vdash e^{*} : tfi}$

    \begin{mathpar}
        \inferrule*[right=\defrule{Local}]{ %% local
            S;(ti^n;l_2;\phi_2) \vdash_i v_l^{*};e^{*} : ti^n;l_2;\phi_2
        } {
            S;C \vdash \<local> \{ i;v_l^{*} \} \; e^{*} \<end> : \epsilon;l_1;\phi_1 \rightarrow ti^n;l_1;\phi_1,\phi_2
        }

        \inferrule*[right=\defrule{Call-Cl}]{ %% call closure
            S \vdash cl : tfi
        } {
            S;C \vdash \<call> cl : tfi
        }

        \inferrule*[right=\defrule{Trap}]{ %% trap
        } {
            S;C \vdash \<trap> : tfi
        }

        \inferrule*[right=\defrule{Label}]{ %% label
            S;C\vdash e_0^{*} : ti_3^{*};l_3;\phi_3 \rightarrow ti_2^{*};l_2;\phi_2 \\
            S;C,\text{label } (ti_3^{*};l_3;\phi_3) \vdash e^{*} : \epsilon;l_1;\phi_1 \rightarrow ti_2^{*};l_2;\phi_2
        } {
            S;C \vdash \<label> \{ e_0^{*} \} \; e^{*} \<end> : \epsilon;l_1;\phi_1 \rightarrow ti_2^{*};l_2;\phi_2
        }
    \end{mathpar}
    \caption{\name Administrative Instruction Rules}
    \label{fig:adminrules}
\end{figure}

Now we will introduce the typing rules for administrative instructions, and the administrative typing judgment in \autoref{fig:adminrules}.
The administrative typing judgment $S;C \vdash e^{*} : tfi$ extends the \name typing rules for instructions to include administrative instructions and the store context $S$.
Every rule of the judgment $C \vdash e^{*} : tfi$ (recall the rules enumerated in \autoref{sec:typesys}) is implicitly added to the administrative judgment by accepting any $S$.

Most of the rules for administrative instructions check against extra information provided by the administrative typing judgment.
\refrule{Local} typechecks a local block using \refrule{Code} to ensure that the body $e^{*}$ is well typed with the indexed types and constraints for local variables provided by the local block as the precondition and any postcondition.
Since local blocks are inline expansions of function calls, we use the optional return postcondition functionality of \refrule{Code} to ensure that returning from inside the local block will be well typed.
\refrule{Call-Cl} typechecks calling a closure by ensuring that the closure $cl$ being called has the same type as the call instruction $\<call> cl$ in $S$.
\refrule{Trap} is always well typed under any precondition and postcondition.
\refrule{Label} typechecks the body of the label block with the precondition of the saved instructions pushed onto the label stack.
If the label was generated by a loop, then the precondition of the saved values is the precondition of the loop, and we know the loop is well typed.
Otherwise, the saved instructions will be an empty sequence and will be well typed from the precondition.

Given these additional typing judgments and rules, we can now show the metatheoretic properties mentioned above.

\section{Relationship Between \wasm and \name}
We want to show two properties about the relationship between \wasm and \name.
First, we want \name to be backwards compatible with \wasm.
It should be possible to convert well-typed \wasm programs into well-typed \name programs with no additional developer effort.
We demonstrate a simple yet naive way of embedding \wasm programs into \name in \autoref{subsec:embedding}.
Second, we want to show that well-typed \name programs can be turned into \wasm programs.
This is accomplished in \autoref{subsec:erasure} using an erasure function that turns \name programs and types into \wasm programs and types.

\input{Chapters/MetaTheory/embedding.tex}
\input{Chapters/MetaTheory/erasure.tex}

\section{Type Safety}
\label{sec:typesafety}
\emph{Type safety} is the property that a well-typed program either reduces to another well-typed program, is an irreducible expression (in the case of \name, a sequence of values), or throws an error (trap, in the case of \name).
Thus, type safety assures us that the behavior of a well-typed program is always well defined.
The type safety of \wasm guarantees a number of important properties, including memory safety.
Proving the type safety of \name gives us a high degree of assurance that it has the same level of safety as \wasm.

\input{Chapters/MetaTheory/subject-reduction.tex}
\input{Chapters/MetaTheory/progress.tex}
