\chapter{Acknowledgements}
\todo{This is pretty out of date, also quite long}
This section proceeds through my life in reverse chronological order.

\section*{University of British Columbia}
First and foremost I would like to thank my wonderful advisors, who made this whole process much easier and significantly contributed to my edification.
\begin{itemize}
    \item William Bowman - Despite being a new professor, William has been a wonderful advisor and provided a large amount of guidance when I was struggling.
    William has been a fantastic resource for all of the work here.
    \item Ivan Beschastnikh - Ivan helped me navigate through the system when I was a wide-eyed first-year and learn from various mistakes I made along the way.
    Ivan was a great source of wisdom and gave me the freedom to pursue the topics I found most interesting.
\end{itemize}

I would also like to thank various professors for providing guidance and advice, some of whom are listed below.
\begin{itemize}
    \item Ron Garcia - Ron is definitely one of the nicest people I have met and has forgotten way more about PL than I have ever learned. He also provided great feedback on this thesis (part of which led to much clearer definitions).
    \item Margo Seltzer - Margo is a kickass professor who is insanely smart and cares a lot about students.
    \item Alan Hu - Alan has provided me with many nuggets of wisdom.
    \item Alex Summers - Alex spent time to give me helpful feedback that greatly improved this thesis.
\end{itemize}

Thank you to Justin Frank, for helping to prove type safety and develop the reference implementation.

Finally, all of my fellow graduate students and lab mates who have become my friends and helped me in various ways. There are way too many awesome people to thank them all, but I include some notable names below.
\begin{itemize}
    \item Anny Gakhokidze, Puneet Mehrotra, Nico Ritschel, Chris Chen, Felipe Ba\~{n}ados Schwerter, Paulette Koronkevich, Aarti Kashyap, Clement Fung, Vaastav Anand, and countless others.
\end{itemize}

%%\section*{University of Washington}
%%During my undergrad, I spent a little over a year working on the Cassius project.
%%This was my first research experience and I think I learned just as much from that experience as the rest of my undergrad combined.
%%\begin{itemize}
%%    \item Pavel Panchekha - Pavel was very patient and provided a lot of help getting me off the ground and learning about what doing research is like.
%%    \item Michael D. Ernst - As well as advising me throughout my undergrad research, Mike guided me through the process of applying to grad school.
%%    \item Zach Tatlock - Zach is super nice and always made sure I wasn't lost.
%%    \item Shoaib Kamil - For all his help in research and helping me find my way at my first conference alone.
%%    \item James Wilcox - For introducing me to PL and helping me find and get started on the Cassius project.
%%    \item All the other profs and graduate students of UWPLSE, who were extremely friendly and welcoming.
%%\end{itemize}

%%\section*{Various Others}
%%Many different people have helped me on my path at various times. Below I list a few notable ones.
%%\begin{itemize}
%%    \item B.J. Unti - For first giving me the idea to go to grad school.
%%    \item Xander Veerhoff - For first teaching me how to program, which was not easy.
%%\end{itemize}

\section*{My Parents}
Last but certainly not least, I'd like to thank my awesome parents for everything they have done and continue to do for me.
I'm not sure one could ask for better parents than I have been given (even if I did not ask for them).
They have both always been there for me, especially my mom who spent a huge amount of effort and time taking care of every issue that came up in my early years, as well as homeschooling me.
My Dad has always been my role model and is the smartest person I know.
He tried to introduce me to set theory when I was around 4 and calculus when I was about 13 (the first one stuck pretty well, the second one not so much).
